New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2022-10-07 paper/paper.Rmd
Mathematica package vs notebook, doublespacing

---
title: |
  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a
    email: ursula.laa@boku.ac.at
  - name: Alex Aumann
    affil: b
    email: aaum0002@student.monash.edu
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
  - name: German Valencia
    affil: b
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      School of Physics and Astronomy, Monash University
  - num: c
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: | 
  This paper describes new user controls for examining high-dimensional data using low-dimensional linear projections and slices. A user can interactively change the contribution of a given variable to a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables. The user can also interactively shift the center of a slice, for example, to explore how structure changes in local subspaces. Two Mathematica notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models. The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls. Some limited implementation has also been made available in the R package \tt{tourr}. 
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

This paper explains the new manual controls for projection and slice tours. The next section describes the new algorithm for manual control, for both projections and slices. The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers. The software section describes a mathematica notebook that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available. The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.

@@

## Mathematica notebook

@@

- Two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.
- Appendix with additional details about manual controls, and Mathematica functions.
- Animated gifs illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

>>>>

---
title: |
  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a
    email: ursula.laa@boku.ac.at
  - name: Alex Aumann
    affil: b
    email: aaum0002@student.monash.edu
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
  - name: German Valencia
    affil: b
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      School of Physics and Astronomy, Monash University
  - num: c
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: | 
  This paper describes new user controls for examining high-dimensional data using low-dimensional linear projections and slices. A user can interactively change the contribution of a given variable to a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables. The user can also interactively shift the center of a slice, for example, to explore how structure changes in local subspaces. The Mathematica package as well as example notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models. The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls. Some limited implementation has also been made available in the R package \tt{tourr}. 
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}\doublespacing
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

This paper explains the new manual controls for projection and slice tours. The next section describes the new algorithm for manual control, for both projections and slices. The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers. The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available. The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the application.

@@

## Mathematica package

@@

- The Mathematica source code defining the new functions in mmtour.wl.
- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
- Appendix with additional details about manual controls, and Mathematica functions.
- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2022-10-07 paper/paper.tex
Mathematica package vs notebook, doublespacing

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{nicematrix}
\NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}

@@

\begin{abstract}
This paper describes new user controls for examining high-dimensional
data using low-dimensional linear projections and slices. A user can
interactively change the contribution of a given variable to a
low-dimensional projection, which is useful for exploring the
sensitivity of structure to particular variables. The user can also
interactively shift the center of a slice, for example, to explore how
structure changes in local subspaces. Two Mathematica notebooks are
provided, which contain functions enabling the user to experiment with
these new manual controls, with one specifically for exploring regions
and boundaries produced by classification models. The advantage of
Mathematica is its linear algebra capabilities, and interactive cursor
location controls. Some limited implementation has also been made
available in the R package \tt{tourr}. 
\end{abstract}

@@

\hypertarget{mathematica-notebook}{%
\subsection{Mathematica notebook}\label{mathematica-notebook}}

@@

\begin{itemize}
\tightlist
\item
  Two Mathematica notebooks, one focused on the application, and another
  that could be used generically for new data.
\item
  Appendix with additional details about manual controls, and
  Mathematica functions.
\item
  Animated gifs illustrating the manual tour and slicing, matching the
  static figures in the paper. These are also available at
  \url{https://vimeo.com/747585410} and
  \url{https://vimeo.com/747590472}.
\end{itemize}

>>>>

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}\doublespacing
\usepackage{graphicx}
\usepackage{nicematrix}
\NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}

@@

\begin{abstract}
This paper describes new user controls for examining high-dimensional
data using low-dimensional linear projections and slices. A user can
interactively change the contribution of a given variable to a
low-dimensional projection, which is useful for exploring the
sensitivity of structure to particular variables. The user can also
interactively shift the center of a slice, for example, to explore how
structure changes in local subspaces. The Mathematica package as well as
example notebooks are provided, which contain functions enabling the
user to experiment with these new manual controls, with one specifically
for exploring regions and boundaries produced by classification models.
The advantage of Mathematica is its linear algebra capabilities, and
interactive cursor location controls. Some limited implementation has
also been made available in the R package \tt{tourr}. 
\end{abstract}

@@

\hypertarget{mathematica-package}{%
\subsection{Mathematica package}\label{mathematica-package}}

@@

\begin{itemize}
\tightlist
\item
  The Mathematica source code defining the new functions in mmtour.wl.
\item
  Three Mathematica notebooks with the code used in the application
  (corresponding to the three subsections).
\item
  Appendix with additional details about manual controls, and
  Mathematica functions.
\item
  Animations illustrating the manual tour and slicing, matching the
  static figures in the paper. These are also available at
  \url{https://vimeo.com/747585410} and
  \url{https://vimeo.com/747590472}.
\end{itemize}
